 

### 2021-07-14 vertical-thermal-review.Rmd
cutting words in T50 section

Temperature can affect photosynthesis via direct and indirect pathways--directly, by altering photosynthetic enzyme activity and electron transport chain, and indirectly through increased $VPD$ causing stomatal closure [@lloydEffectsRisingTemperatures2008]. Photosynthesis has a peaked response to temperature, with the peak commonly found to correspond to the prevalent ambient growing season conditions [Tan et al. 2017; @slotSituTemperatureResponse2017]. Beyond the optimum, photosynthesis decreases as a result of stomatal closure (e.g. @slotSituTemperatureResponse2017; Smith et al. 2020 [temperate/boreal refs needed), and eventually due to biochemical constraints (@kumarathungeAcclimationAdaptationComponents2019a; Sage & Kubien 2007; Varhammer et al. 2015). Irresversible damage to photosystem II, leading to leaf necrosis and eventually leaf death occurs at $T_{leaf}$ of $40-60^{\circ}C$
[Baker 2008; @feeley_thermal_2020].

>>>>

Temperature can affect photosynthesis via direct and indirect pathways--directly, by altering photosynthetic enzyme activity and electron transport chain, and indirectly through increased $VPD$ causing stomatal closure [@lloydEffectsRisingTemperatures2008]. Photosynthesis has a peaked response to temperature, with the peak commonly found to correspond to the prevalent ambient growing season conditions [@tanOptimumAirTemperature2017; @slotSituTemperatureResponse2017]. Beyond the optimum, photosynthesis decreases as a result of stomatal closure (e.g. @slotSituTemperatureResponse2017; Smith et al. 2020 [temperate/boreal refs needed), and eventually due to biochemical constraints (@kumarathungeAcclimationAdaptationComponents2019a; Sage & Kubien 2007; Varhammer et al. 2015). Irresversible damage to photosystem II, leading to leaf necrosis and eventually leaf death occurs at $T_{leaf}$ of $40-60^{\circ}C$
[Baker 2008; @feeley_thermal_2020].
